---
title       : Introduction to R
subtitle    : Vectors in R
author      : Adam J Sullivan
job         : 
license     : by-nc-nd
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, quiz, bootstrap, interactive]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
logo        : shield_image.png
biglogo     : shield_image_large.png
knit        : slidify::knit2slides
assets      : {assets: ../../assets}
---  .segue bg:grey

# Vectors in R


--- .class #id

## What is a Vector?

- The fundamental data type in R is the vector. 
- A vector is a sequence of data elements of the same type. 
 


--- .class #id 

## Creating Vectors
 
What we have used here is the concatenation operator which takes the arguments and places them in a vector in the order in which they were entered. 

```{r}
 
 x <- c(1,5,2, 6)
 x
 
 is.vector(x)
```
 

--- .class #id

##  Vector Arithmetic
 
- We can do arithmetic with vectors in a similar manner as we have with integers. 
- When we use operators we are doing something element by element or "elementwise." 
 
```{r}
 y <- c(1,6,4,8)
 x+y
```


--- .class #id

## Elementwise

It is important to remember what happens when we consider an ``elementwise" operation
 
```{r}
 
 x*y
 x/y
x %% y
```

--- .class #id
 
 
##  Recycling

- We do have to be careful when performing arithmetic operations on vectors. 
- There is a concept called recycling and this happens when 2 vectors do not have the same length

--- .class #id

## Recycling Example
 
```{r}
 
 z<- c(1,2 ,6 ,8, 9, 10)
 
 x+z
```

--- .class #id

##  Recycling

- Intuition would make us think that we could not perform this operation when the length of both vectors are not the same. 
- However what R does is it rewrites `x` such that we have `x <- c(1 , 5, 2, 6 , 1, 5)`.
- This is called recycling, when R makes the shorter vector longer by repeating elements in the order they are listed in.

--- .class #id

## Recyling
 
```{r}
x+ z
c(1 , 5, 2, 6 , 1, 5) + z
```
 
--- .class #id

##  Functions on Vectors
 
- There are various functions that we can run over a vector and as we continue on we will learn more about these functions.
 
- One of the simplest functions can help us with knowing information about Recycling that we encountered before. This is the `length()` function. 

--- .class #id

## Functions on Vectors

```{r}
 
length(x)
length(y)
length(z)
```

--- .class #id

## Functions on Vectors


- Then length vector is very important with the writing of functions which we will get to in a later unit.  
 
- We can use ***any()*** and ***all()*** in order to answer logical questions about elements 
 
--- .class #id

## Functions on Vectors

```{r}
 
any(x>3)
all(x>3)
 
```
--- .class #id

##  Built in Functions

 There area various other functions that can be run on vectors some you have seen in other classes.
 
- ***mean()*** finds the arithmetic mean of a vector. 
- ***median()*** finds the median of a vector.
- ***sd()*** and ***var()*** finds the standard deviation and variance of a vector respectively.

--- .class #id

##  Built in Functions

- ***min()*** and ***max()*** finds the minimum and maximum of a vector respectively.
- ***sort()*** returns a vector that is sorted. 
- ***summary()*** returns a 5 number summary of the numbers in a vector. 

--- .class #id

##  `which()` Function
 
- Some functions help us work with the data more to return values in which we are interested in. 
- For example, above we asked if any elements in vector `x` were greater than 3. 
- The `which()` function will tell us the elements that are. 
 
```{r}
 which(x>3)
```

--- .class #id

##  Vector Indexing
 
We can call specific elements of a vector by using the following:
 

- `x[]` is a way to call up a specific element of a vector.
- `x[1]` is the first element.
- `x[3]` is the third element.
- `x[-3]` is a vector with everything but the third element. 

--- .class #id
 
##  Working with Vectors 
 
```{r}
 #List elements to make sure we have what we need
 ls()
 
 x[3]
 
 x[-3]
```

--- .class #id

## Replacing Values

- We have seen how to subtract an element from a vector but we can use the same information to place it back in. 
- We start with the same vector `x` that we started with. 
 
```{r}
 x
 x<-x[-3]
 x
 
```

--- .class #id

## Insering Values

 We can then add the original element back in
 
```{r}
 x <- c(x[1:2], 2, x[3])
 x
 
```

--- .class #id

##  Indexing with Booleans

- Before we used `any(x > 3)` and `which(x > 3)`. 
- Now we can see not only their position in the vector, but indexing allows us to return their values. 
 
```{r}
 x[x > 3]
```

--- .class #id

##  Creating Vectors
 
 There are multiple ways we can create a vector but we must let R know what we are doing
 
```{r}
 
 y[1] <- 3
 y[2] <- 15
 
```

--- .class #id

##  Creating Vectors
 
 Within R, we have not defined any `y` yet so it will not create a vector in this manner. There are multiple ways of creating vectors:
 
```{r}
 
 y1 <- vector(length=2)
 y1[1] <- 3
 y1[2] <- 15
 
 y1
```

--- .class #id

## Creating Vectors

```{R}
 
 y2 <-  c(3,15)
 y2
 
 y3 <- seq(from=3, to=15, length=2)
 y3
 
 y4 <- seq(from=3, to=15, by=12)
 y4
```
 
--- .class #id

## Creating Vectors

 Aside from these ways to create the specific vector of (3,15) we can create vectors a couple more ways as well
 
```{r}
 y5 <- 3:10
 y5
 
 y6 <- rep(8,6)
 y6
 
```

--- .class #id

## Creating Vectors

```{R} 
 y7 <- rep(c(1,2,3),3)
 y7
 
```

--- .class #id

##  Naming Vector Elements

- With vectors it can be important to assign names to the values. 
- Then when doing plots or considering maximum and minimums, instead of being given a numerical place within the vector we can be given a specific name of what that value represents. 
- For example say that vector `x` represents the number of medications of 4 unique patients. We could then use the ***name()*** function to assign names to the values

--- .class #id

```{r}
x
names(x)

names(x) <- c("Patient A", "Patient B", "Patient C", "Patient D")
x
```