---
title       : Introduction to R
subtitle    : Matrices in R
author      : Adam J Sullivan
job         : 
license     : by-nc-nd
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, quiz, bootstrap, interactive]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
logo        : shield_image.png
biglogo     : shield_image_large.png
knit        : slidify::knit2slides
assets      : {assets: ../../assets}
---  .segue bg:grey

# Lists

--- .class #id

## Lists

- Within R a list is a structure that can combine objects of different types. 
- We will learn how to create and work with lists in this section. 

--- .class #id

## Creating Lists

- A list is actually a vector but it does differ in comparison to the other types of vectors which we have been using in this class.
    - Other vectors are *atomic vectors*
    - A list is a type of vector called a *recursive vector*.

--- .class #id

## An Example Database

We first consider a patient database where we want to store their

- Name
- Amount of bill due
- A Boolean indicator of whether or not they have insurance. 

--- .class #id

## Types of Information

We then have 3 types of information here: 

- character
- numerical
- logical. 

--- .class #id

## Single Patient

To create a list of one patient we say
```{r}
a <- list(name="Angela", owed="75", insurance=TRUE)
a
```

--- .class #id

## Indexing
 
- Notice that unlike a typical vector this prints out in multiple parts. 
- This also allows us to help with indexing as we will see below. 
- There is another easy way to create this same list

--- .class #id

## Creating the Same List
 
```{r} 
 a.alt <- vector(mode="list")
 a.alt[["name"]] <- "Angela"
 a.alt[["owed"]] <- 75
 a.alt[["insurance"]] <- TRUE
 
 a.alt
```

--- .class #id

## Lists of Lists
 
- We could then create a list like this for all of our patients. 
- Our database would then be a list of all of these individual lists.

--- .class #id
 
## List Operations 

 
- With vectors, arrays and matrices, there was really only one way to index them. 
- However with lists there are multiple ways:

--- .class #id

## List Indexing

```{r}
a[["name"]]
a[[1]]
a$name
```

--- .class #id

## Double vs Single Brackets

- All of the previous are ways to index data in a list. 
- Notice that in two of the above we used double brackets. 
- Next we see the difference between double and single brackets.

--- .class #id

## Double vs Single Brackets

```{r}
a[1]

class(a[1])

a[[1]]
class(a[[1]])
```

--- .class #id

## Double vs Single Brackets

- With the single bracket we are  returned another list.
- With the double bracket we are returned an element in the original class of what kind of data we entered. 
- Depending on your goals you may want to use single or double brackets. 

--- .class #id

## Adding and Subtracting Elements

With a list we can always add more information to it. 
```{r}
a$age <- 27 
a
```

--- .class #id

## Adding and Subtracting Elements

In order to delete an element from a list we set it to NULL. 

```{r}
a$owed <- NULL
a
```

--- .class #id


## List Components and Values

In order to know what kind of information is included in a list we can look at the ***names()*** function
```{r}
names(a)
```

--- .class #id

## Unlisting

To find the values of things we could go ahead and unlist them
```{r}
a.un <- unlist(a)

a.un
class(a.un)
```

--- .class #id

## Unlisting

- If There is Character data in the original list that unlisted everything will be in character format. 
- If your list contained all numerical elements than the class would be numerical. 

--- .class #id

## Applying Functions to Lists

- Just like arrays and matrices we can use an ***apply()*** function. 
- Specifically we have ***lapply()*** and ***sapply()*** functions for lists. 
- With the original ***apply()*** function we could specify whether the function was applied to either the rows or the columns. 
- With the case of lists both functions are applied to elements of the list.

--- .class #id

## Applying Functions to Lists

```{r}
#Number list
n <- list(1:5, 6:37)
n
```

--- .class #id

## Applying Functions to Lists

- With this list we see that we have two separate vectors of numbers included. 
- Then let us see the results of either using ***lapply()*** and ***sapply()***

--- .class #id

## Applying Functions to Lists

```{r}
lapply(n, median)

sapply(n, median)

```

--- .class #id

## Apply Functions an Lists


- The ***lapply()*** function returns a list with the median of each of the original lists. 
- While the ***sapply()*** function returns a vector of the medians. 

--- .class #id

## Recursive Lists

- Before it was mentioned that a list is a recursive vector. 
- This is because we can actually have lists within lists. 

--- .class #id

## Recursive Lists

For example let us go back to our patient data. 
```{r}
s <- list(name="Chandra", insurance="TRUE", age=36)

patients <- list(a,s)
patients

```

--- .class #id


## Final Notes on Lists

- It is important to remember how we can call these features of lists. 
- Many of you will want to use R for model building and regressions. 
- You almost never want to use the generated output from R. 
- For example R does not automatically return the confidence intervals with a regression. 

--- .class #id

## Final Notes on Lists

- The output from most regression functions in R is actually a list. 
- What this means is I can extract the elements from the list that I want in order to build tables that display the exact information that I want it to. 
- This is why we take the time to discuss how to search what is in a list and how to access it. 

--- .class #id

##  Example with Output of a List

```{r} 
x <- rnorm(500,10, 3)
y <- 3*x + rnorm(500, 0, 2)


```

--- .class #id

##  Example with Output of a List

```{r} 

fit <- lm(y~x)
fit
```

--- .class #id

##  Example with Output of a List

- So R just gave me the coefficients back but no other information. 
- This means my knowledge of accessing lists is key. 

```{r}
names(fit)
```

--- .class #id

##  Example with Output of a List

- I can see that R actually has a lot more information that they did not display for me. 
- Next I consider a function where it summarizes the information from this model

--- .class #id

##  Example with Output of a List

```{r}
summary <- summary(fit)
summary
```

--- .class #id

##  Example with Output of a List

```{r}
names(summary)
```

--- .class #id

## Conclusion of Lists

- R has so much information about regression that is never even displayed unless I dig deeper. 
- Understanding lists and accessing information means you can output custom tables that look much more professional than what R gives you. 