---
title       : Introduction to R
subtitle    : Dataframes in R
author      : Adam J Sullivan
job         : 
license     : by-nc-nd
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, quiz, bootstrap, interactive]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
logo        : shield_image.png
biglogo     : shield_image_large.png
knit        : slidify::knit2slides
assets      : {assets: ../../assets}
--- .segue bg:grey

# DataFrames in R



--- .class #id



## Dataframe

- With statistics we are most likely to use the data structure called a data frame. 
- This is similar to a matrix in appearance however we can have multiple types of data in it like a list.
- Each column must contain the same type of data or R will most likely default to character for that column. 
- It is very important that you become proficient in working with data frames in order to fully understand data analysis.

--- .class #id

## Creating Data Frames
We usually create a data frame with vectors. 

```{r} 
names <- c("Angela", "Shondra")
ages <- c(27,36)
insurance <- c(TRUE, T)
patients <- data.frame(names, ages, insurance)
patients
```

--- .class #id




## Adding Rows or Columns

- We may wish to add rows or columns to our data. 
- We can do this with:
  - ***rbind()*** 
  - ***cbind()*** 

--- .class #id

## Adding Rows or Columns

For example we can go back to our patient data and say we wish to add another patient we could just do the following
```{r}
l <- c(names="Liu Jie", age=45, insurance=TRUE)
rbind(patients, l)
```

>- This warning serves as a reminder to always know what your data type is.
>- R has read our data in as a factor when we want it as a character. 

--- .class #id

## Adding Rows or Columns

```{r}
patients$names <- as.character(patients$names)
patients <- rbind(patients, l)
patients
```

--- .class #id

## Adding Rows or Columns

Finally if we decided to then place another column of data in we could
```{r}

# Next appointments
next.appt <- c("09/23/2016", "04/14/2016", "02/25/2016")

#Lets R know these are dates
next.appt <- as.Date(next.appt, "%m/%d/%Y")
next.appt
```

--- .class #id

```{r}
patients <- cbind(patients, next.appt)
patients
```

--- .class #id

## Accessing Data Frames

In order to best consider accessing of data frames we will use some built in data from R. 

```{r}
library(datasets)
titanic <- data.frame(Titanic)
```

--- .class #id

## Variables Included in Titanic

```{r}
colnames(titanic)
```

--- .class #id

## Preview Into Data


```{r} 
titanic[1:2,]

head(titanic)
```


--- .class #id

## Indexing 

- Indexing is the same as that for matrices. 
- ***head()*** function allows us to access the first rows of the data frame. 
- We can also access data by using both column and row information

--- .class #id

## Indexing

```{r}
# accessing age information
titanic[,3]

#accessing age information using column name
titanic[, "Age"]

```


--- .class #id

## Indexing and Naming

- This means we can access data with a column or row number
- More importantly we can use the name. 
- For large data frames accessing by a name is key.

--- .class #id


## Further Indexing 

Say we want to know information about a particular class
```{r}
titanic["1st", ]

```

--- .class #id

## Further Indexing

We could also ask for information by using the factors that we have as well

```{r}
first.class.freq <- titanic[titanic$Class=="1st", "Freq"]
first.class.freq


male.freq <- titanic[titanic$Sex=="Male", "Freq"]
male.freq
```

--- .class #id

## Our New Variables

```{r}
sum(first.class.freq)
sum(male.freq)
```

--- .class #id

## Adding New Variables 

- Suppose we not only want to know the frequency of survival but the proportion 
- We can ask R to calculate this and add it to our data. 

```{r}
titanic$surv_p <- titanic$Freq/sum(titanic$Freq)
head(titanic,4)
```

--- .class #id

## Replacing Values 

- Perhaps we were not pleased the decimal places and want to have this as a percentage. 
- We can overwrite the values and change this. 

```{r}
titanic$surv_p <- titanic$surv_p*100
head(titanic,4)
```

