---
title       : Introduction to R
subtitle    : Tibbles in R
author      : Adam J Sullivan
job         : 
license     : by-nc-nd
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, quiz, bootstrap, interactive]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
logo        : shield_image.png
biglogo     : shield_image_large.png
knit        : slidify::knit2slides
assets      : {assets: ../../assets}
--- .segue bg:grey

# Tibbles in R



--- .class #id



```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
set.seed(4561)
```



## Tibbles

Previosly we have worked with data in the form of

- Vectors
- Lists
- Arrays
- Dataframes

--- .class #id


## Tibbles

- *"Tibbles"* are a new modern data frame.  
- It keeps many important features of the original data frame. 
- It removes many of the outdated features. 

--- .class #id

## Compared to Data Frames

- A *tibble* never changes the input type. 
    - No more worry of characters being automatically turned into strings. 
- A tibble can have columns that are lists.
- A tibble can have non-standard variable names.
    - can start with a number or contain spaces. 
    - To use this refer to these in a backtick.
- It only recyles vectors of length 1. 
- It never creates row names.

--- .class #id


## Column-Lists

```{r, message=FALSE}
library(tidyverse)
try <- tibble(x = 1:3, y = list(1:5, 1:10, 1:20))
try 


#try <- as_data_frame(c(x = 1:3, y = list(1:5, 1:10, 1:20)))
#try
# Leads to error
```

--- .class #id


## Non-Standard Names

```{R}
names(data.frame(`crazy name` = 1))
names(tibble(`crazy name` = 1))
```

--- .class #id

## Coercing into Tibbles

- A tibble can be made by coercing `as_tibble()`. 
- This works similar to `as.data.frame()`.
- It works efficiently.

--- .class #id


## Coercing into Tibbles


```{r}
l <- replicate(26, sample(100), simplify = FALSE)
names(l) <- letters

microbenchmark::microbenchmark(
  as_tibble(l),
  as.data.frame(l)
)
```

--- .class #id

## Tibbles vs Data Frames

There are a couple key differences between tibbles and data frames. 

- Printing.
- Subsetting. 

--- .class #id

## Printing

- Tibbles only print the first 10 rows and all the columns that fit on a screen. - Each column displays its data type. 
- You will not accidently print too much. 


```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```

--- .class #id



## Subsetting



- We can index a tibble in the manners we are used to
    - `df$x`
    - `df[["x"]]`
    - `df[[1]]`
- We can also use a `pipe` which we will learn about later.
    - `df %>% .$x`
    - `df %>% .[["x"]]`

--- .class #id

## Subsetting

```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df$x
df[["x"]]
df[[1]]
```

--- .class #id

## Subsetting

```{r}
df %>% .$x
df %>% .[["x"]]
df %>% .[[1]]
```